---
title: "Empirisk analyse"
author: "Simon"
date: "3/6/2022"
output: html_document
---

```{r}
library(tidyverse)
```


```{r}
library(readxl)
c_20 <- read_excel("C:/Users/Simon ik mig/Desktop/9. Semester/Projekt/Data/c_20.xlsx")
c_20


data= c_20 %>% 
  select(Closingprice, Date) %>%
  mutate(month = months(Date), year= format(Date,format="%Y"))
```

Create volatility for each month as the deviation from the mean in each month, so the cumulative squared deviation will be calculated for each month. Using the closing price. 


```{r}

means=data %>%
  group_by(year, month) %>%
  summarise(mean_month = mean(Closingprice)) 
```
We now got the means, we can bind it to the dataset using month and year

```{r}
options(scipen = 999)

per_dev=data %>%
  left_join(means, by= c("month", "year"))%>%
  mutate(percent_dev= round((Closingprice-mean_month)/mean_month*100,2))%>%
  mutate(percent_dev_abs= round(abs((Closingprice-mean_month)/mean_month*100),2)) %>%
  mutate(percent_dev_sq= round(((Closingprice-mean_month)/mean_month*100)^2,2)) 
```

Lastly we can take the mean percentage deviation over the month. 

```{r}

vol_mes=per_dev %>%
  group_by(year, month) %>%
  summarise(mean_abs= mean(percent_dev_abs), mean_sq=mean(percent_dev_sq)) %>%
  mutate(month = ifelse(month == "januar", "01", ifelse(month == "februar", "02", ifelse(month == "marts", "03", ifelse(month == "april", "04", ifelse(month == "maj", "05", ifelse(month == "juni", "06",ifelse(month == "juli", "07", ifelse(month == "august", "08", ifelse(month == "september", "09", ifelse(month == "oktober", "10", ifelse(month == "november", "11", "12")))))))))))) %>%
  mutate(day= "01") %>%
  mutate(date= as.Date(ISOdate(year = year, month = month, day = day)))  %>%
  ungroup()%>%
  select(mean_abs, mean_sq, date)%>%
  arrange(date)

```

creating time series 

```{r}
mean_sq= ts(vol_mes$mean_sq, start = c(1997,1), frequency = 12)

plot(mean_sq)

mean_abs= ts(vol_mes$mean_abs, start = c(1997,1), frequency = 12)

plot(mean_abs)
```



## Sandard deviation

```{r}
sd_data=data %>%
  group_by(year, month) %>%
  summarise(sd= sd(Closingprice)) %>%
  mutate(month = ifelse(month == "januar", "01", ifelse(month == "februar", "02", ifelse(month == "marts", "03", ifelse(month == "april", "04", ifelse(month == "maj", "05", ifelse(month == "juni", "06",ifelse(month == "juli", "07", ifelse(month == "august", "08", ifelse(month == "september", "09", ifelse(month == "oktober", "10", ifelse(month == "november", "11", "12")))))))))))) %>%
  mutate(day= "01") %>%
  mutate(date= as.Date(ISOdate(year = year, month = month, day = day)))  %>%
  ungroup()%>%
  select(sd, date)%>%
  arrange(date)
  
```

plotting

```{r}
sd= ts(sd_data$sd, start = c(1997,1), frequency = 12)
plot(sd)
```


# BBD

```{r}
library(readxl)
US_MPU_Monthly <- read_excel("C:/Users/Simon ik mig/Desktop/9. Semester/Projekt/Data/US_MPU_Monthly.xlsx")

```
```{r}
BBD_data= US_MPU_Monthly %>%
  mutate(day= "01") %>%
  mutate(date= as.Date(ISOdate(year = Year, month = Month, day = day)))  %>%
  select(-c(Year, Month, day))%>%
  arrange(date)
```
```{r}
BBD_AWN= ts(BBD_data$`BBD MPU Index Based on Access World News`, start= c(1985,1), frequency = 12)
BBD_10MP= ts(BBD_data$`BBD MPU Index Based on 10 Major Papers`, start = c(1985,1), frequency = 12)


plot(BBD_AWN)
plot(BBD_10MP)
```

# VIX measure

```{r}
library(readr)
VIX_2004 <- read_csv("C:/Users/Simon ik mig/Desktop/9. Semester/Projekt/Data/VIX_2004.csv")
```

```{r}
data_vix= VIX_2004 %>%
  mutate(date = as.Date(DATE, format = "%m/%d/%Y" ))%>%
  mutate(month = format(date,"%m"), year= format(date,format="%Y"))%>%
  select(CLOSE, month, year) %>%
  group_by(year, month)%>%
  summarise(mean_vix= mean(CLOSE))
  
```

```{r}
VIX= ts(data_vix$mean_vix, start = c(1990,1), frequency = 12)
plot(VIX)
```


# ARCH

```{r}
library("forecast")
library("rugarch")
library("aTSA")
library("urca")
```


```{r}
c_20= c_20$Closingprice

plot(log(c_20))

Acf(c_20)

Pacf(c_20)

auto.arima(c_20,d = 0, D = 0, max.p = 5,max.q = 5, max.d = 0,max.D = 0, stepwise = TRUE,   seasonal = FALSE)

model_est <- arima(c_20, order = c(0,0,5), method = "ML")

test_lb <- Box.test(model_est$residuals, lag = 10, type = c("Ljung-Box"), fitdf = 2)
print(test_lb)

Acf(model_est$residuals, lag.max=10)

Acf(model_est$residuals^2, lag.max=10)

test_arch <- arch.test(model_est, output = FALSE)
print(test_arch)

spec1 <- ugarchspec(variance.model=list(model="sGARCH"),
                 mean.model=list(armaOrder=c(2,4))) ## Så det her er ARCH-M

fit1 <- ugarchfit(c_20,spec=spec1)
show(fit1)


spec2 <- ugarchspec(variance.model=list(model="sGARCH"),
                 mean.model=list(armaOrder=c(2,4),archm=TRUE))

fit2 <- ugarchfit(c_20,spec=spec2)
show(fit2)
```

